Program Transformations for miniKanren

### 2020-05-22 papers/miniKanren/text/nonconj.tex
add pseudocode algo-1
add some refs to fig 1

In this section we will describe a novel approach to specialization of relational programs.
This approach draws inspiration from both conjunctive partial deduction and supercompilation.
The aim was to create a specialization algorithm which is simpler than conjunctive partial deduction and uses properties of \mk{} to improve performance of the input programs.

The high-level idea behind the algorithm is to select a relation call to unfold by using a heuristic which decides if the call can narrow down the answer set.
A driving tree is constructed for the selected call in isolation.
The leaves of the computed tree are examined.
If all leaves are either computed substitutions or are calls to some relations accompanied with non-empty substitutions, then the leaves are collected and are put back into the root conjunction instead of the examined call.
According to denotational semantics of \mk{} it is safe to compute individual conjuncts in any order, thus it is ok to drive any call and then propagate its results onto the other calls.

>>>>

\begin{figure}[!t]
  \centering
  \includegraphics[width=0.7\textwidth]{algo_pseudo-crop.pdf}
  \caption{\db{Non-conjunctive Partial Deduction Pseudo Code}}
  \label{fig:ncpd-pseudo}
\end{figure}

In this section we will describe a novel approach to specialization of relational programs.
This approach draws inspiration from both conjunctive partial deduction and supercompilation.
The aim was to create a specialization algorithm which is simpler than conjunctive partial deduction and uses properties of \mk{} to improve performance of the input programs.
\db{The algorithm pseudocode is shown on Fig.~\ref{fig:ncpd-pseudo}\footnote{
    \db{For the sake of breavity and clarity, we provide functions \texttt{drive\_disj} and \texttt{drive\_conj} describing driving processes for disjunction and conjunction correspondly. Driving itself if a trivial combination of presented functions (line 2)}}}.

The high-level idea behind the algorithm is to select a relation call to unfold by using a heuristic which decides if the call can narrow down the answer set \db{(line 15)}.
A driving tree is constructed for the selected call in isolation \db{(line 16)}.
The leaves of the computed tree are examined.
If all leaves are either computed substitutions or are calls to some relations accompanied with non-empty substitutions, then the leaves are collected and are put back into the root conjunction instead of the examined call \db{(lines 19,20)}.
According to denotational semantics of \mk{} it is safe to compute individual conjuncts in any order, thus it is ok to drive any call and then propagate its results onto the other calls.
